Score-based tests of measurement invariance: Use in practice

### 2013-10-22 frontiers/frontiers.Rnw
update differentiation for weighted and max statistics. Add motivation for simulation. Delete figure panels showing zero powers.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@279 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\section{Simulation 1}

In Simulation 1, we examined the extent to which the proposed tests
are sensitive to the specific model parameter that violates measurement
invariance.  If, say, a factor loading violates measurement
invariance, it is plausible that this violation impacts other
parameter estimates, including factor covariances or other loadings 
associated with the factor in question.  Thus, the goal of the
Simulation 1 is to examine the extent to which the proposed tests
attribute the measurement invariance violation to the parameters that
are truly in violation.

@@

Full simulation results are presented in Figure~\ref{fig:sim11res} to 
Figure~\ref{fig:sim13res}; 
Figure~\ref{fig:sim11res} displays power curves as a function of 
violation magnitude for the first
factor loading $\lambda_{11}$, with panels reflecting the parameters 
being tested and lines reflecting different statistics.
Figures~\ref{fig:sim12res} and Figure~\ref{fig:sim13res} display 
similar power curves when the factor covariance 
$\phi_{12}$ and error variance $\epsilon_{11}$ violate measurement
invariance, respectively. 
All three test statistics exhibit
similar results, with $\max\mathit{LM}_{uo}$ demonstrating lower power across all situations.  As described earlier, this statistic 
is asymptotically equivalent to the LRT.  $\mathit{WDM}_o$  and $\max \mathit{LM}_o$ exhibited equivalent power curves because, when only one parameter is tested, the statistics are equivalent.

From these figures, one generally observes that the tests isolate 
the parameter violating measurement invariance.  Additionally, the tests 
have somewhat higher power to detect measurement invariance violations 
in the factor loading and factor covariance parameters, as opposed to the error variance
parameter.  Finally, simultaneous tests of all factor loadings or all error
parameters result in decreased power, as compared to the situation
where one tests only the violating parameter.  This occurs because, in
testing a subset of parameters (only one of which violates measurement
invariance), we are effectively dampening the signal of a measurement
invariance violation.

@@

\begin{figure}
\caption{Simulated power curves for $\max \mathit{LM}_o$ test, $\mathit{WDM}_o$ test, and
$\mathit{LM}_{uo}$ test across three levels of the ordinal variable $m$ and measurement invariance 
violations of 0--4 standard errors (scaled by $\sqrt{n}$). The parameter violating measurement 
invariance is $\lambda_{11}$.}
\label{fig:sim11res}
<<sim11res, fig=TRUE, height=9, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
       subset = (parms == "loading" & 
		 pars %in% c("Loading1", "All Loadings", "All Errors") & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude", ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	       strip.default(which.given, factor.levels = parlabs[c(1, 4:5)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

@@

\begin{figure}
\caption{Simulated power curves for $\max \mathit{LM}_o$ test, $\mathit{WDM}_o$ test, and
$\mathit{LM}_{uo}$ across three levels of the ordinal variable $m$, and measurement invariance
violations of 0--4 standard errors (scaled by $\sqrt{n}$). The parameter violating measurement
invariance is $\phi_{12}$.}
\label{fig:sim12res}
<<sim12res, fig=TRUE, height=9, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
       subset = (parms == "var" & 
		 pars %in% c("Covariance", "All Loadings", "All Errors") & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude", ylab="Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	       strip.default(which.given, factor.levels = parlabs[c(2, 4:5)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

@@

\begin{figure}
\caption{Simulated power curves for $\max \mathit{LM}_o$ test, $\mathit{WDM}_o$ test, and $\mathit{LM}_{uo}$ test across three levels of the ordinal variable $m$ and measurement invariance 
violations of 0--4 standard errors (scaled by $\sqrt{n}$). The parameter violating measurement invariance is $\psi{11}$}
\label{fig:sim13res}
<<sim13res, fig=TRUE, height=9, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
       subset = (parms == "error" & 
		 pars %in% c("Error1", "All Loadings", "All Errors") & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude", ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	       strip.default(which.given, factor.levels = parlabs[c(3,4:5)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

@@

\begin{figure}
\caption{Simulated power curves for $\max \mathit{LM}_o$ test, $\mathit{WDM}_o$ test, and $\mathit{LM}_{uo}$ test across three levels of the ordinal variable $m$ and measurement invariance 
violations of 0--4 standard errors (scaled by $\sqrt{n}$). The parameter violating measurement invariance is the unmodeled loading.}
\label{fig:sim21res}
<<sim21res, fig=TRUE, height=9, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra" & 
		 pars %in% c("Loading1", "Error1", "Covariance") & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude", ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	       strip.default(which.given, factor.levels = parlabs[c(1, 2:3)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

>>>>

\section{Simulation 1}
Factor analysis models(and SEMs) are usually the first step for social scientists to determine the latent construct. These models are fairly complex with many inter-related parameters. Previous applications of the proposed tests have typically involved simpler models. In this paper we intended to study whether the tests can still identify the specific parameters that violate measurement invariance when parameter correlation is present in the model. \seeme{the above part is added motivation. Please check.} Particularly, 
in Simulation 1, we examined the extent to which the proposed tests
are sensitive to the specific model parameter that violates measurement
invariance.  If, say, a factor loading violates measurement
invariance, it is plausible that this violation impacts other
parameter estimates, including factor covariances or other loadings 
associated with the factor in question.  Thus, the goal of the
Simulation 1 is to examine the extent to which the proposed tests
attribute the measurement invariance violation to the parameters that
are truly in violation.

@@

Full simulation results are presented in Figure~\ref{fig:sim11res} to 
Figure~\ref{fig:sim13res}; 
Figure~\ref{fig:sim11res} displays power curves as a function of 
violation magnitude for the first
factor loading $\lambda_{11}$, with panels reflecting the parameters 
being tested and lines reflecting different statistics.
Figures~\ref{fig:sim12res} and Figure~\ref{fig:sim13res} display 
similar power curves when the factor covariance 
$\phi_{12}$ and error variance $\epsilon_{11}$ violate measurement
invariance, respectively. In those Figures, we only show test parameters demonstrating non-zero powers for clarity. \seeme{Add this sentence to explaining there is no zero power curves.}
All three test statistics exhibit
similar results, with $\max\mathit{LM}_{uo}$ demonstrating lower power across all situations.  As described earlier, this statistic 
is asymptotically equivalent to the LRT.  $\mathit{WDM}_o$  and $\max \mathit{LM}_o$ exhibited equivalent power curves because, when only one parameter is tested, the statistics are equivalent.

From these figures, one generally observes that the tests isolate 
the parameter violating measurement invariance.  Additionally, the tests 
have somewhat higher power to detect measurement invariance violations 
in the factor loading and factor covariance parameters, as opposed to the error variance
parameter.  Finally, simultaneous tests of all factor loadings or all error
parameters result in decreased power, as compared to the situation
where one tests only the violating parameter.  This occurs because, in
testing a subset of parameters (only one of which violates measurement
invariance), we are effectively dampening the signal of a measurement
invariance violation. This ''dampening'' effect is more apparent for $\max \mathit{LM}_o$ statistic, because it sums across all parameters tested and then take the maximum. Whereas $\mathit{WDM}_o$ takes the maximum over parameters being tested directly.\seeme{Add from ''This'' to ''directly''.Please check.}

@@

\begin{figure}
\caption{Simulated power curves for $\max \mathit{LM}_o$ test, $\mathit{WDM}_o$ test, and
$\mathit{LM}_{uo}$ test across three levels of the ordinal variable $m$ and measurement invariance 
violations of 0--4 standard errors (scaled by $\sqrt{n}$). The parameter violating measurement 
invariance is $\lambda_{11}$.}
\label{fig:sim11res}
<<sim11res, fig=TRUE, height=9, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
       subset = (parms == "loading" & 
		 pars %in% c("Loading1", "All Loadings") & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude", ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	       strip.default(which.given, factor.levels = parlabs[c(1, 4)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

@@

\begin{figure}
\caption{Simulated power curves for $\max \mathit{LM}_o$ test, $\mathit{WDM}_o$ test, and
$\mathit{LM}_{uo}$ across three levels of the ordinal variable $m$, and measurement invariance
violations of 0--4 standard errors (scaled by $\sqrt{n}$). The parameter violating measurement
invariance is $\phi_{12}$.}
\label{fig:sim12res}
<<sim12res, fig=TRUE, height=9, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
       subset = (parms == "var" & 
		 pars %in% c("Covariance") & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude", ylab="Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	       strip.default(which.given, factor.levels = parlabs[c(2)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

@@

\begin{figure}
\caption{Simulated power curves for $\max \mathit{LM}_o$ test, $\mathit{WDM}_o$ test, and $\mathit{LM}_{uo}$ test across three levels of the ordinal variable $m$ and measurement invariance 
violations of 0--4 standard errors (scaled by $\sqrt{n}$). The parameter violating measurement invariance is $\psi{11}$}
\label{fig:sim13res}
<<sim13res, fig=TRUE, height=9, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
       subset = (parms == "error" & 
		 pars %in% c("Error1", "All Errors") & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude", ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	       strip.default(which.given, factor.levels = parlabs[c(3,5)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

@@

\begin{figure}
\caption{Simulated power curves for $\max \mathit{LM}_o$ test, $\mathit{WDM}_o$ test, and $\mathit{LM}_{uo}$ test across three levels of the ordinal variable $m$ and measurement invariance 
violations of 0--4 standard errors (scaled by $\sqrt{n}$). The parameter violating measurement invariance is the unmodeled loading.}
\label{fig:sim21res}
<<sim21res, fig=TRUE, height=9, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra" & 
		 pars %in% c("Loading1", "Error1") & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude", ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	       strip.default(which.given, factor.levels = parlabs[c(1, 2:3)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@
